Identification of Direct Cherenkov Pixels using Boosted Decision Trees

### 2016-07-01 report/draft1.tex
Backup 02/07/16

\documentclass[11pt]{article}
\usepackage{graphicx}
\graphicspath{ {graphs/} }
\usepackage{geometry}
\geometry{a4paper, portrait, margin=1.0in}
\usepackage{gensymb}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{pgfplotstable, booktabs}

@@

To fully reconstruct an event, we will need to find the charge, energy and the core position coordinates. In addition, we will need to determine the height of the first atmospheric interaction. We would normally require at least five data points in order to reconstruct these variables. However, it is possible to use the non-observation of light in one telescope to constrain the core position. Thus, we require that an event was be seen by at least four telescopes for a five-telescope array such as HESS.

\section{DC Pixel Identification}
Cosmic Rays will emit Direct Cherenkov (DC) light in the upper atmosphere, before generating an Extended Air Shower (EAS) through interaction with the lower atmosphere. In telescope images, the DC light is usually concentrated in a single  \textquoteleft DC pixel'. Identifying this pixel is challenging, because the brighter EAS Cherenkov light background often overlaps with the DC pixel. In order to apply the LPD method to data, we must first identify the DC pixel in a shower image, and determine the number of DC photoelectrons present.

\subsection{Theoretical LPD}
To understand why DC light has a characteristic distribution, it is important to recall that Cherenkov emission begins when a charged particle passes the local speed of light in a medium.  We know that the refractive index of air is proportional to the density of air. Thus, the minimum velocity needed for Cherenkov Emission is lower for low atmospheric altitudes, because the air is denser. If the velocity is sufficient, we will observe photon emission from the particle at the Cherenkov Angle $\theta_{c}$.

@@

We obviously require that the Cherenkov Angle is real, so then require $\cos(\theta_{c}) \leq 1$. The minimum required velocity $\beta_{min}$ is thus equal to the inverse of the local refractive index $\eta$. The velocity of a particle is a function of its energy, determined with the relations $\gamma = \frac{E}{m_{0} c^{2}}$ and $\beta^{2} = 1 - \frac{1}{\gamma ^{2}}$. Thus, we find that the minimum velocity ultimately leads to an Energy per unit mass threshold. If we assume that the slight difference in proton and neutron mass is negligible here, than we can determine a local energy per nucleon threshold that will be valid for all Cosmic Rays passing a given atmospheric altitude.

The Energy per Nucleon Threshold for Cherenkov Light Emission as a function of height is illustrated in \ref{fig:generalenergy}. Once the Energy of a Cosmic Ray exceeds the local Cherenkov Energy Threshold of the atmosphere, the Nucleus will begin emitting a ring of Cherenkov Light.  Then for a given Telescope Array altitude above sea level and a given height h, simple trigonometry yields the final radius of emission on the ground:
\[ Radius(h) = \tan (\theta_{C}(h)) \times (h - altitude_{array})\]

@@

We see that there is a clear characteristic photon distribution which, for energies above 25 TeV, is almost constant except for a varying endpoint. However, we will not be able to directly measure this LPD. Instead, a telescope images containing air showers must be analysed, and from these images, we can infer the quantity of DC light present amongst a large degree of background. Extensive simulations of Camera images will be performed, in order to quantify the resultant LPD extracted from the telescope images. After determining a likely experimental LPD, Monte Carlo events can be simulated, and then reconstructed.

\subsection{Image Simulation}
We can find the DC pixel by defining the variable $Q_{DC} = \frac{Intensity_{N.N.max}}{Intensity}$ as the ratio of the largest neighbouring pixel intensity to the intensity of a given pixel. In previous experiments, the DC pixel candidate was identified by applying a number of cuts to pixels in an image \cite{hess07}, and from the subset of pixels passing the specified cuts, selecting the pixel with the smallest $Q_{DC}$ as the \textquoteleft DC candidate'. Due to the low pass rate for cuts, we obtain a small low-contamination dataset, while the majority of telescope images are left without a DC pixel candidate. An improved method would aim to increase the number of correctly identified DC pixels, while still enabling cuts which discriminate well between correctly and incorrectly identified DC pixels.

Classifiers provide an alternative method of identification, making use of supervised machine learning to find rules for categorising pixels. To train a classifier, we require a set of training pixels, as well as the correct class for each pixel. Once trained, a classifier can then be used to predict the class of a pixel.

The CORSIKA package \cite{Heck98} was used to generate Cosmic Ray events, and the Sim\textunderscore telarray package  \cite{Bernlohr08} was used to generate corresponding HESS array telescope images.  Simulation with EAS background was used to produce training pixel sets, while corresponding simulation without EAS background was used to find the true DC pixel in each training image. As part of a new method developed for this analysis, a Boosted Decision Tree (BDT) classifier was trained with the data to identify DC pixels. Once trained, the BDT was applied to a separate \textquoteleft testing' set of simulated telescope images. As before, the true DC pixel was determined from a second EAS-free simulation. Thus, the accuracy of BDT identification for test telescope images was calculated.

The full simulation of air showers was performed using the CORSIKA package with a standard atmospheric profile derived from measurements conducted at the HESS site in Namibia. This Atmospheric Profile is included in the Appendix. In total, 5000 training events and a further 10000 testing events were simulated. The simulated particles were $Fe^{56}$, within the Energy Range of $35-135$ TeV and a flux spectrum $\phi(E) \propto E^{-2.7}$. For each set of simulated event, 4 unique random number seeds were used to generate the shower. An altitude of 1800m was assumed, again corresponding to the HESS site. The simulated zenith angle ranged from $0\degree < \theta < 2 \degree$, while the simulated azimuth angle ranges from $-2\degree < \phi < 2 \degree$. The four smaller HESS-phase-1 telescopes were arranged in a cross along the x/y axis with the larger HESS-phase-2 \textquoteleft CT5' telescope placed at the center. The length of each cross arm was $85m$. The simulated target region of the cores was chosen to be a square centered on CT5, with each 300m-long side bisecting the x/y axis.Due to hardware differences between CT5 and the original HESS 1 telescopes, we analyse HESS1 and HESS2 images separately. To determine the true class of each pixel, a simulation was initially run with an energy cut of 10 PeV on all muons and electrons. Because this cut exceeded the primary particle energy, neither daughter muons and electrons, nor the photons they would have emitted, were simulated. Thus the hadronic Cherenkov Light from the primary particle and daughter fragments, but not the EAS light, was present in the camera image. A second identical \textquoteleft EAS Simulation' was run including the same random seeds, but without the energy cut on muons and electrons. This gave a complete EAS image including identical DC light.

@@

\begin{figure}
\newgeometry{a4paper, portrait, margin=1.5in}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[trim=80 120 80 150,clip,width=\textwidth]{graphDC}
\caption{A typical camera image without the EAS shower. The DC light is visible in every telescope, indicated by the white arrow. The DC pixel is circled in white. The largest telescope image is from CT5, but was not used in analysis.}
\label{fig:DCtelimage}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[trim=80 120 80 150,clip,width=\textwidth]{graphfull}
\caption{The same shower as in \ref{fig:DCtelimage} is shown here with the inclusion of the EAS shower. The DC light is pixels indicated with a white arrow and circle. The shower center of gravity is marked by a white cross.}
\label{fig:cutdistribution2}
\end{minipage}
\restoregeometry
\end{figure}

@@

\subsection{Classic HESS1 DC Pixel Identification}
As a basis for comparison, the original HESS cuts listed in Table \ref{tab:table1} were replicated for the set of test data. For every HESS1 image, the total image amplitude $I_{tot}$ was used alongside the zenith angle $\theta$ to determine a dynamic cut, $Q_{DC} < 0.14 \times \log(\frac{I_{tot}}{161 \times \cos \theta})$. Among those pixels passing all cuts, the one with the smallest $Q_{DC}$ was selected as the DC pixel candidate for the image. Because many images had no pixel that passed all cuts, the $Q_{DC}$ method was frequently unable to identify a DC pixel. In the original analysis, an additional cut $r_{core} \textgreater 40m$ was applied. However, the uncertainty in determining the core position through Hillas Analysis is typically of the order of $\pm 30m$. Consequently, this particular cut was omitted.

\begin{table}[h!]
  \centering
  \caption{Cuts applied to image pixel sets, used by HESS collaboration \cite{hess07}}
  \label{tab:table1}
  \begin{tabular}{ccc}
    \toprule
    Variable & Cut\\
    \midrule
     $ \Delta_{C.o.G}$ & \textgreater 0.17 \\
     $ \Delta_{C.o.G}$ & \textless 0.91 \\
     $\Delta_{Direction}$ & \textless 0.45 \\
     $\Delta_{Line}$ & \textless 0.23 \\
     Aspect Ratio & \textless 0.75 \\
     $Q_{DC}$ & \textless 0.14$ \times \log(\frac{I_{tot}}{161 \times \cos \theta})$ \\
    \bottomrule
  \end{tabular}
\end{table}

The candidates were checked against the true DC pixels identified in the EAS-free images. From the testing sample, 1.81\% of all HESS1 images were correctly identified and passed the cuts. Once misidentified events were considered, the post-cuts sample was 77.8\% accurate, as shown in Figure \ref{fig:cutdistribution}. These values will serve as a useful benchmark for BDT performance. Increasing the acceptance rate will be essential for enabling spectroscopic analysis of cosmic rays, by providing more events to analyse.

@@

\subsection{Boosted Decision Tree DC Identification}
As an alternative to use of $Q_{DC}$, a new method of DC pixel identification was developed as part of this analysis by using a BDT trained with the Scikit Learn Python package \cite{scikit-learn}. Using the training set of 2000 CORSIKA events, and randomly split further, with 90\% in a learning subset and 10\% in a subset to check for overtraining. Within the learning subset, every HESS 1 image was used, provided it was triggered in both EAS-free and full-shower simulations. For each of the triggered image pixels, an entry was formed with the variables listed in table \ref{tab:hess1classifier}. A class of 0 was assigned to every non-DC pixel, and a class of 1 was assigned to every DC pixel. Having created a dataset, the BDT was then trained with a maximum depth of 8, and with 100 trees generated. The data was provided in the form of individual pixel entries, rather than as discrete sets for images or events.

\begin{table}[h!]
  \centering
  \caption{Relative Feature Importance in HESS-1 BDT training}
  \label{tab:hess1classifier}
  \begin{tabular}{ccc}
    \toprule
    Variable & Relative Importance\\
    \midrule
    $DC_{Count}$ & 0.33\\
    $Mean_{N.N}$ & 0.24\\
    $Q_{DC}$ & 0.14\\
    $\Delta_{Direction}$ & 0.12\\
    Image Amplitude & 0.07\\
    $raw_{Q}$ & 0.05\\
    $\Delta_{Line}$ & 0.02\\
    $Intensity$ & 0.02\\
    \bottomrule
  \end{tabular}
\end{table}

The relative importance of each \textquoteleft feature' is automatically calculated by the Scikit Learn package, and is also recorded in table \ref{tab:hess1classifier}. The variable $DC_{Count}$ was consistently the most importance variable across many combinations of included variables and BDT training parameters. It was found that, under the conditions listed above, the BDT was 99.94 \% accurate for the learning pixel subset, and 99.94 \%  accurate for the overtraining-check pixel subset. This indicates that the BDT was not significantly overtrained, which would otherwise be manifested by a large divergence in accuracy between learning and overtraining-check data.

Having trained the BDT successfully, it was then applied to the same test dataset as for the classic $Q_{DC}$ identification. In each camera image, the event with the largest BDT score was deemed to be \textquoteleft most signal-like', and thus selected as the DC pixel candidate. A cut was applied, requiring $P_{signal} > 0.5$ for the DC candidate to be accepted. A second cut requiring $DC_{Count} > 150$ removed many incorrectly identified events. Application of this combined cut greatly increases the successful identification rate. From the testing sample, 16.01\% of all images were correctly identified and passed the cuts. The BDT was found to be 79.2\% accurate in identifying DC pixels which passed the cuts. This represents a very significant improvement in pixel identification efficiency, as well as a minor increase in accuracy after cuts.

@@

\begin{table}[h!]
  \centering
  \caption{Comparison of $Q_{DC}$ and HESS 1 BDT Performance}
  \label{tab:qdcbdtcomparison1}
  \begin{tabular}{ccc}
    \toprule
    & $Q_{DC}$ & BDT\\
    \midrule
    Pixels Accepted and Correctly Identified (\%)  & 1.81 & 16.01\\
   Pixels Accepted and Incorrectly Identified (\%)  & 0.52& 4.22\\
    Sample Purity (\%)  & 77.8 & 79.2 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{HESS-2 BDT}
The original HESS study predated the construction of the larger CT5 telescope, and thus focused exclusively in the four HESS-1 telescopes. Although the cuts in \ref{tab:qdccuts} were not optimised for the differing pixel size and angular viewing region of CT5, they were replicated to provide a basis for comparison. Based on CT5 images from the training sample, an increased 5.6\% of DC pixels were correctly identified and passed the required cut. A further 11.3 \% of all pixels passed the cuts, despite being misidentified. This represented a very heavy decrease in sample purity to just 33.0\%. We can assume that better optimised CT5 cuts could remove many of these misidentified events, though it is unlikely that any significant improvement in the number of correctly identified DC pixels would be possible. Thus, the number of DC pixels successfully identified using the $Q_{DC}$ method is still a valid benchmark for comparative BDT performance. A true accuracy rate closer to the HESS 1 rate of 90\% could be expected. Due to the distinctiveness of the CT5 telescope, direct application of the HESS1 classifier to the CT5 telescope yields poorer BDT performance. To improve CT5 pixel identification, a separate HESS2 BDT was instead trained with the same variables as above. The relative feature importance is listed in Table \ref{tab:hess2classifier}.

@@

For the training events, the classifier was 99.97\% accurate, while for the overtraining check data, it was 99.96\% accurate. The classifier was thus not significantly overtrained. Applying the new classifier lede to an improvement in CT5 classification, with 17.9 \% of pixels being correctly identified and accepted, and an accuracy of 66.1\%. Despite this improvement, a clear gap emerged between the between the performance of HESS1 classifier on old telescopes, and the performance of the HESS2 classifier on CT5. Part of this discrepancy can be explained by the HESS2 learning dataset, which has just one DC pixel per event rather than the four HESS1 DC pixels. Additionally, HESS 2 images will have roughly 2600 non-DC pixels per event, versus 3600 DC pixels per event for HESS 1 images. The HESS2 classifier was thus trained on significantly fewer DC pixels, so we would expect to see the clear performance gap in comparison to the HESS1 classifier.

\begin{table}[h!]
  \centering
  \caption{Comparison of $Q_{DC}$ and HESS 2 BDT Performance}
  \label{tab:qdcbdtcomparison2}
  \begin{tabular}{ccc}
    \toprule
    & $Q_{DC}$ & BDT\\
    \midrule
    Pixels Accepted and Correctly Identified (\%) & 5.59 & 18.52 \\
   Pixels Accepted and Incorrectly Identified (\%) & 14.75 & 4.44 \\
    Sample Purity (\%) & 27.5 & 80.6 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{High Multiplicity Events}
For LPD event reconstruction, we require at least four measurements of the DC light. We define the telescope multiplicity of an event as the number of DC pixels which are identified and pass all cuts. Telescope multiplicity measures both HESS1 and HESS2 pixels. A \textquoteleft high-multiplicity event' is one in which we have telescope multiplicity $>3$. If we only consider high-multiplicity events, we can determine the $Q_{DC}$ performance for use in LPD reconstruction. After applying this cut, there were was just one single high-multiplicity event found using the $Q_{DC}$ method. This meant that just 0.01\% of all pixels were accepted, although with a sample purity of 100\%. Because the final sample consisted of just three HESS1 pixels, and one HESS2 pixel, the associated Poissonian error in acceptance rates and accuracy rates will be extremely large.

>>>>

\documentclass[11pt]{article}
\usepackage{graphicx}
\graphicspath{ {graphs/} }
\usepackage{geometry}
\geometry{a4paper, portrait, margin=1.0in}
\usepackage{gensymb}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{pgfplotstable, booktabs}
\usepackage{caption}
\usepackage{subcaption}

@@

To fully reconstruct an event, we will need to find the charge, energy and the core position coordinates. In addition, we will need to determine the height of the first atmospheric interaction. We would normally require at least five data points in order to reconstruct these variables. However, it is possible to use the non-observation of light in one telescope to constrain the core position. Thus, we require that an event was be seen by at least four telescopes for a five-telescope array such as HESS. An illustration of the HESS array is shown in Figure \ref{fig:hessdiagram}.

\begin{figure}
\begin{center}
\includegraphics[height=0.3\textheight]{hessdiagram}
\caption{A diagram illustrating the HESS array \cite{hess07}. The four small telescopes are shown, but not the larger central telescope. The smaller Cherenkov ring is seen alongside the broader EAS light. A simplified camera image is also shown, and can be compared to a full image seen in Figure \ref{fig:fulltelimage}.}
\label{fig:hessdiagram}
\end{center}
\end{figure}

\section{DC Pixel Identification}
In a Comsic Ray event, the \textquoteleft Primary Particle' will emit Direct Cherenkov (DC) light in the upper atmosphere, before generating an Extended Air Shower (EAS) through interaction with the lower atmosphere. The primary particle is the Cosmic Ray itself, while the EAS refers to all daughter particle of the primary particle. In telescope images, the DC light is usually concentrated in a single  \textquoteleft DC pixel'. Identifying this pixel is challenging, because the brighter EAS Cherenkov light background often overlaps with the DC pixel. In order to apply the LPD method to data, we must first identify the DC pixel in a shower image, and determine the number of DC photoelectrons present.

\subsection{Theoretical LPD}
To understand why DC light has a characteristic distribution, it is important to recall that Cherenkov emission begins when a charged particle passes the local speed of light in a medium.  We know that the refractive index of air is proportional to the density of air. Thus, the minimum velocity needed for Cherenkov Emission is lower for low atmospheric altitudes, because the air is denser. If the velocity is sufficient, we will observe photon emission from the particle at the Cherenkov Angle $\theta_{c}$. We obviously require that the Cherenkov Angle is real, so then require $\cos(\theta_{c}) \leq 1$. The minimum required velocity $\beta_{min}$ is thus equal to the inverse of the local refractive index $\eta$.

@@

The velocity of a particle is a function of its energy, determined with the relations $\gamma = \frac{E}{m_{0} c^{2}}$ and $\beta^{2} = 1 - \frac{1}{\gamma ^{2}}$. Thus, we find that the minimum velocity ultimately leads to an Energy per unit mass threshold. If we assume that the slight difference in proton and neutron mass is negligible here, than any atmospheric altitude will have a local Energy per Nucleon threshold, and this threshold will be the same for all Cosmic Rays.

The Energy per Nucleon Threshold for Cherenkov Light Emission as a function of height is illustrated in Figure \ref{fig:generalenergy}. Once the Energy of a Cosmic Ray exceeds the local Cherenkov Energy Threshold of the atmosphere, the Nucleus will begin emitting a ring of Cherenkov Light.  Then, for a given Telescope Array altitude above sea level and a given height h, simple trigonometry yields the final radius of emission on the ground:
\[ Radius(h) = \tan (\theta_{C}(h)) \times (h - altitude_{array})\]

@@

For energies above 25 TeV, we see that there is a clear characteristic photon distribution which is almost independent of Energy except for a varying endpoint. The LPD displays many discontinuities, which can be explained by the interpolation of Refractive Index. Exponential Interpolation will not provide a function that is smooth for second or third derivatives, and consequently, the LPD is not smooth. In reality, there will be a funtion describing the Refractive Index as a function of height, but it will depend on many variables such as Temperature. This will be the dominant cause of variations to the LPD, and thus, uncertainty in the LPD from Refractive Index Interpolation will be sub-dominant.

In any case, we will not be able to directly measure this LPD. Instead, a telescope images containing air showers must be analysed, and from these images, we can infer the quantity of DC light present amongst a large degree of background. Extensive simulations of Camera images will be performed, in order to quantify the resultant LPD extracted from the telescope images. After determining a likely experimental LPD, Monte Carlo events can be simulated, and then reconstructed.

\subsection{Image Simulation}
The CORSIKA package \cite{Heck98} was used to generate Cosmic Ray events, and the Sim\textunderscore telarray package  \cite{Bernlohr08} was used to generate corresponding HESS array telescope images.  Simulation with EAS background was used to produce training pixel sets, while corresponding simulation without EAS background was used to find the true DC pixel in each training image.

The full simulation of air showers was performed using CORSIKA with a standard atmospheric profile derived from measurements conducted at the HESS site in Namibia. This is included in the Appendix. In total, 5000 training events and a further 10000 testing events were simulated. The simulated particles were $Fe^{56}$, within the Energy Range of $35-135$ TeV and a flux spectrum $\phi(E) \propto E^{-2.7}$. For each set of simulated event, 4 unique random number seeds were used to generate the shower. An altitude of 1800m was assumed, again corresponding to the HESS site. The simulated zenith angle ranged from $0\degree < \theta < 2 \degree$, while the simulated azimuth angle ranges from $-2\degree < \phi < 2 \degree$. The four smaller HESS-phase-1 telescopes were arranged in a cross along the x/y axis with the larger HESS-phase-2 \textquoteleft CT5' telescope placed at the center. The length of each cross arm was $85m$. The simulated target region of the cores was chosen to be a square centered on CT5, with each 300m-long side bisecting the x/y axis. Due to hardware differences between CT5 and the original HESS 1 telescopes, we analyse HESS1 and HESS2 images separately.

To determine the true class of each pixel, a simulation was initially run with an energy cut of 10 PeV on all muons and electrons. Because this cut exceeded the primary particle energy, neither daughter muons and electrons, nor the photons they would have emitted, were simulated. Thus the hadronic Cherenkov Light from the primary particle and daughter fragments, but not the EAS light, was present in the camera image. A second identical \textquoteleft EAS Simulation' was run including the same random seeds, but without the energy cut on muons and electrons. This gave a complete EAS image including identical DC light. The difference is well-illustrated in Figures \ref{fig:DCtelimage} and \ref{fig:fulltelimage}.

@@

\begin{figure}
\newgeometry{a4paper, portrait, margin=1.0in}
\centering
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[trim=80 120 80 150,clip,width=\textwidth]{graphDC}
\caption{An ideal camera image without the EAS shower. The DC light is visible in every telescope, indicated by the white arrow. The DC pixel is circled in white. The largest telescope image is from CT5, and is analysed separately.}
\label{fig:DCtelimage}
\end{minipage}\hfill
\begin{minipage}{0.45\textwidth}
\centering
\includegraphics[trim=80 120 80 150,clip,width=\textwidth]{graphfull}
\caption{This is a typical camera image. The same shower as in Figure \ref{fig:DCtelimage} is shown here with the inclusion of the EAS shower.  The DC light is pixels indicated with a white arrow and circle. The shower center of gravity is marked by a white cross.}
\label{fig:fulltelimage}
\end{minipage}
\restoregeometry
\end{figure}

@@

\subsection{Classic HESS1 DC Pixel Identification}
We can find the DC pixel in am image by defining the variable $Q_{DC} = \frac{Intensity_{N.N.max}}{Intensity}$ as the ratio of the largest neighbouring pixel intensity to the intensity of a given pixel. In previous experiments, the DC pixel candidate was identified by applying a number of cuts to pixels in an image \cite{hess07}, and from the subset of pixels passing the specified cuts, selecting the pixel with the smallest $Q_{DC}$ as the \textquoteleft DC candidate'. Due to the low pass rate for cuts, we obtain a small low-contamination dataset, while the majority of telescope images are left without a DC pixel candidate.

As a basis for comparison, these original HESS cuts listed in Table \ref{tab:qdccuts} were replicated for the set of test data. For every HESS1 image, the total image amplitude $I_{tot}$ was used alongside the zenith angle $\theta$ to determine a dynamic cut, $Q_{DC} < 0.14 \times \log(\frac{I_{tot}}{161 \times \cos \theta})$. Because many images had no pixel that passed all cuts, the $Q_{DC}$ method was frequently unable to identify a DC pixel. In the original analysis, an additional cut $r_{core} \textgreater 40m$ was applied. However, the uncertainty in determining the core position through Hillas Analysis is typically of the order of $\pm 30m$. Consequently, this particular cut was omitted.

\begin{table}[h!]
  \centering
  \caption{Cuts applied to image pixel sets, used by HESS collaboration \cite{hess07}}
  \label{tab:qdccuts}
  \begin{tabular}{ccc}
    \toprule
    Variable & Cut\\
    \midrule
     $ \Delta_{C.o.G}$ & \textgreater 0.17 \\
     $ \Delta_{C.o.G}$ & \textless 0.91 \\
     $\Delta_{Direction}$ & \textless 0.45 \\
     $\Delta_{Line}$ & \textless 0.23 \\
     Aspect Ratio & \textless 0.75 \\
     $Q_{DC}$ & \textless 0.14$ \times \log(\frac{I_{tot}}{161 \times \cos \theta})$ \\
    \bottomrule
  \end{tabular}
\end{table}

The candidates were checked against the true DC pixels identified in the EAS-free images. From the testing sample, 1.81\% of all HESS1 images were correctly identified and passed the cuts. Once misidentified events were considered, the post-cuts sample was 77.8\% accurate, as shown in Figure \ref{fig:cutdistribution}. These values will serve as a useful benchmark for BDT performance. Increasing the acceptance rate will be essential for enabling spectroscopic analysis of cosmic rays, by providing more events to analyse. An improved method would aim to increase the number of correctly identified DC pixels, while still enabling cuts which discriminate well between correctly and incorrectly identified DC pixels.

@@

\subsection{Boosted Decision Tree DC Identification}
Classifiers provide an alternative method of DC pixel identification, making use of supervised machine learning to find rules for categorising pixels. To train a classifier, we require a set of training pixels, as well as the correct class for each pixel. Once trained, a classifier can then be used to predict the class of a pixel. As part of a new method developed for this analysis, a Boosted Decision Tree (BDT) classifier was trained to identify DC pixels using the Scikit Learn Python package \cite{scikit-learn}.

A Decision Tree is a simple method of classification, in which a number of branching decisions are used to place events into one of many \textquoteleft leaves'. A leaf lies at the end of every decision chain, and will be associated with one of the possible classes. Any event being assigned to a particular leaf will then be classified as belonging to that leaf's class. Each branch and leaf has a sample, which indicates the number of training events that are placed in there. As part of the machine learning process, branches are constructed to maximise leaf purity. A hypothetical perfect branch would split a mixed sample into two pure leaves.

It will almost always be possible to construct a tree, such that each event was placed in its own, unique leaf. The performance of a tree would be perfect for the training set. However, if applied to a separate testing sample, the performance would be significantly worse. The tree would not be classifying data, but rather simply remembering the structure of the training data. We describe such a tree as being overtrained.

In general, larger training sets prevent overtraining by making it more difficult to uniquely distinguish the individual events. The maximum depth of a tree is the number of branches a decision chain can pass through before it must end on a leaf. The minimum leaf sample is the minimum number that every leaf has sample must exceed. Either of these quantities can be used to restrict overtraining of a Decision Tree.

A graphical representation of a Decision Tree is given in Figure \ref{fig:decisiontree}. Decision Tree training is highly unpredictable, meaning similar datasets can produce wildly differing trees. Thus Boosted Decision Trees are used to provide a more stable method of classification. The training of a Boosted Decision Tree is in some sense an averaging process, where many different Decision Trees are trained. One starting Decision Tree is produced, with each event having an equal weighting. After each iteration of tree production, any misidentified events in a tree are given a higher weighting for the training of the next tree \cite{Roe:2004na}. Over time, this leads to an improved learning process by focussing more on those events that are more difficult to classify.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{decisiontreehess1}
\caption{An example of a decision tree for the HESS 1 telescope system. The left direction is followed if the branch condition is true, while the right direction is followed if the condition is false. Red leaves contain non-DC pixels, while blue leaves represent DC pixels. The sample of each leaf is listed. Darker boxes have purer samples, while lighter boxes are more mixed.}
\label{fig:decisiontree}
\end{center}
\end{figure}

In this case, the training set consisted of pixels from 5000 CORSIKA events. It was randomly split further, with 90\% in a learning subset and 10\% in a subset to check for overtraining. Within the learning subset, every HESS 1 image was used, provided it was triggered in both EAS-free and full-shower simulations. For each of the triggered image pixels, an entry was formed with the variables listed in table \ref{tab:hess1classifier}. A class of 0 was assigned to every non-DC pixel, and a class of 1 was assigned to every DC pixel. Having created a dataset, the BDT was then trained with a maximum depth of 8, and with 1000 trees generated. The data was provided in the form of individual pixel entries, rather than as discrete sets for images or events.

\begin{table}[h!]
  \centering
  \caption{Relative Feature Importance in HESS-1 BDT training}
  \label{tab:hess1classifier}
  \begin{tabular}{ccc}
    \toprule
    Variable & Relative Importance\\
    \midrule
    $DC_{Signal}$ & 0.33\\
    $Mean_{N.N}$ & 0.24\\
    $Q_{DC}$ & 0.14\\
    $\Delta_{Direction}$ & 0.12\\
    Image Amplitude & 0.07\\
    $raw_{Q}$ & 0.05\\
    $\Delta_{Line}$ & 0.02\\
    $Intensity$ & 0.02\\
    \bottomrule
  \end{tabular}
\end{table}

The relative importance of each \textquoteleft feature' is automatically calculated by the Scikit Learn package, and is also recorded in table \ref{tab:hess1classifier}. The variable $DC_{Signal}$ was consistently the most importance variable across many combinations of included variables and BDT training parameters. It was found that, under the conditions listed above, the BDT was 99.94 \% accurate for the learning pixel subset, and 99.94 \%  accurate for the overtraining-check pixel subset. This indicates that the BDT was not significantly overtrained, which would otherwise be manifested by a large divergence in accuracy between learning and overtraining-check data.

Having trained the BDT successfully, it was then applied to the same test dataset as for the classic $Q_{DC}$ identification. In each camera image, the event with the largest BDT score was deemed to be \textquoteleft most signal-like', and thus selected as the DC pixel candidate. A cut was applied, requiring $P_{signal} > 0.5$ for the DC candidate to be accepted. A second cut requiring $DC_{Signal} > 150$ removed many incorrectly identified events. Application of this combined cut greatly increases the successful identification rate. From the testing sample, 16.01\% of all images were correctly identified and passed the cuts. The BDT was found to be 79.2\% accurate in identifying DC pixels which passed the cuts. This represents a very significant improvement in pixel identification efficiency, as well as a minor increase in accuracy after cuts.

@@

\begin{table}[h!]
  \centering
  \caption{Comparison of $Q_{DC}$ and HESS 1 BDT Performance}
  \label{tab:qdcbdtcomparison1}
  \begin{tabular}{ccc}
    \toprule
    & $Q_{DC}$ & BDT\\
    \midrule
    Accepted Pixels Correctly Identified (\%)  & 1.81 & 16.01\\
   Accepted Pixels Incorrectly Identified (\%)  & 0.52& 4.22\\
    Sample Purity (\%)  & 77.8 & 79.2 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{HESS-2 BDT}
The original HESS study predated the construction of the larger CT5 telescope, and thus focused exclusively in the four HESS-1 telescopes. Although the cuts in Table \ref{tab:qdccuts} were not optimised for the differing pixel size and angular viewing region of CT5, they were replicated to provide a basis for comparison. Based on CT5 images from the training sample, an increased 5.6\% of DC pixels were correctly identified and passed the required cut. A further 11.3 \% of all pixels passed the cuts, despite being misidentified. This represented a very heavy decrease in sample purity to just 33.0\%. We can assume that better optimised CT5 cuts could remove many of these misidentified events, though it is unlikely that any significant improvement in the number of correctly identified DC pixels would be possible. Thus, the number of DC pixels successfully identified using the $Q_{DC}$ method is still a valid benchmark for comparative BDT performance. A true accuracy rate closer to the HESS 1 rate of 90\% could be expected, at the expense of a depressed acceptance rate. Due to the distinctiveness of the CT5 telescope hardware, direct application of the HESS1 classifier to the CT5 telescope yields poorer BDT performance. Discrepancies in To improve CT5 pixel identification, a separate HESS2 BDT was instead trained with the same variables as above. The relative feature importance is listed in Table \ref{tab:hess2classifier}.

@@

For the training events, the classifier was 99.97\% accurate, while for the overtraining check data, it was 99.96\% accurate. The classifier was thus not significantly overtrained. Applying the new classifier lede to an improvement in CT5 classification, with 18.52 \% of pixels being correctly identified and accepted, and an accuracy of 80.6\%. With this improvement, there is a no significant gap between the between the performance of HESS1 classifier on old telescopes, and the performance of the HESS2 classifier on CT5.

\begin{table}[h!]
  \centering
  \caption{Comparison of $Q_{DC}$ and HESS 2 BDT Performance}
  \label{tab:qdcbdtcomparison2}
  \begin{tabular}{ccc}
    \toprule
    & $Q_{DC}$ & BDT\\
    \midrule
    Accepted Pixels Correctly Identified (\%) & 5.59 & 18.52 \\
    Accepted Pixels Incorrectly Identified (\%) & 14.75 & 4.44 \\
    Sample Purity (\%) & 27.5 & 80.6 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{High Multiplicity Events}
For LPD event reconstruction, we require at least four measurements of the DC light. We define the telescope multiplicity of an event as the number of DC pixels which are identified and pass all cuts. Telescope multiplicity measures both HESS1 and HESS2 pixels. A \textquoteleft high-multiplicity event' is one in which we have telescope multiplicity $>3$. If we only consider high-multiplicity-$Q_{DC}$ events, we can determine how frequently the $Q_{DC}$ method provides events suitable for LPD reconstruction. After applying the multiplicity cut, there were was just one single high-multiplicity event found using the $Q_{DC}$ method out of the 10000 testing events. This meant that just 0.01\% of all pixels were accepted, albeit with a sample purity of 100\%. Because the final sample consisted of just three HESS1 pixels, and one HESS2 pixel, the associated Poissonian error in acceptance rates and accuracy rates will be extremely large.

Through application of the two HESS classifiers, we can find the comparative high-multiplicity-BDT performance. If we only consider high-multiplicity-BDT events, the fraction of correctly identified HESS 1 DC pixels falls to 2.20\%, while the fraction of incorrectly identified pixels passing the cuts falls to 0.43\%. The sample purity increases slightly to 83.8\%.  In total 1.61\% of the HESS 2 pixels are correctly identified and accepted, although the accuracy rate is increased to 88.0\%. The relatively high fraction of passing events is in excess of the random expectation of $0.16^{4}=0.07\%$ and $0.19^{4}=0.13\%$ for HESS1 and HESS2. This suggests that DC pixel identification between different telescope images is strongly correlated. The same is true for the $Q_{DC}$ high multiplicity rates.

The minor discrepancy in performance between HESS1 and HESS2 can partially be explained by the geometry of the HESS array. Any event  with a core position near the central CT5 telescope is likely to trigger the four outer telescopes. However, the dim inner LPD is unlikely to be seen in the big central telescope. In the event of a core position near one of the corner telescopes, CT5 and at least three of the other telescopes are likely to have a clear DC signal. Thus the effective area for high multiplicity events including CT5 is smaller than for those involving the HESS1 telescopes.

Despite the minor performance difference, the BDT method is clearly superior in both cases. On top the gains in BDT accuracy when considering only high-multiplicity events high multiplicity events, we also find the relative performance gap over the $Q_{DC}$ method is vastly increased. As a result of the two-hundred-fold increase in acceptance rate for high-multiplicity-BDT events against high-multiplicity-$Q_{DC}$ events, the expected data sample size will increase dramatically. This should enable spectroscopic analysis of Cosmic Ray elements to be conducted after a much shorter experimental run time that for the $Q_{DC}$ method. Use of the BDT identification method will be assumed throughout the rest of this paper, and a high-multiplicity event will be intended to mean a high-multiplicity-BDT event.
